---
title: "A Comparative Analysis of Valence, Duration, and Energy in the Music of JJ Lin, Joker Xue, and Leehom Wang"
author: 
  - John Zhang
  - Heyucheng Zhang
  - Ariel Xing
  - Haoan Wang
thanks: "Code and data are available at: https://github.com/Clearsky21z/Spotify_Analysis"
date: today
date-format: long
abstract: "This study examines the trends in valence, duration, and energy in the music of three prominent Chinese pop artists: JJ Lin, Joker Xue, and Leehom Wang. Valence, representing the emotional tone of a song, generally shows a slight decline across all three artists, with Joker Xue experiencing the most significant drop. Song duration remains stable for Joker Xue and Leehom Wang but increases for JJ Lin over time. In contrast, energy trends show JJ Lin and Joker Xue with declining intensity, while Leehom Wang's music becomes more energetic."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(usethis)
library(spotifyr)
library(ggplot2)
library(knitr)

jj_lin <- readRDS(here("data/01-raw_data/jj_lin.rds"))
leehom_wang <- readRDS(here("data/01-raw_data/leehom_wang.rds"))
joker_xue <- readRDS(here("data/01-raw_data/joker_xue.rds"))

# Convert to tibble
jj_lin <- as_tibble(jj_lin)
leehom_wang <- as_tibble(leehom_wang)
joker_xue <- as_tibble(joker_xue)
```

# Introduction

The evolution of musical trends in popular music is a well-documented phenomenon, often reflecting changes in cultural preferences, artistic expression, and technological advancements. In recent years, the global music scene has seen significant contributions from Chinese pop artists, among whom JJ Lin, Joker Xue, and Leehom Wang stand out. Each of these artists has built a reputation for their unique contributions to Mandarin pop music, characterized by distinct emotional tones and production qualities.

Valence, energy, and song duration are key musical attributes that influence how audiences perceive songs. Valence represents the positivity or emotional tone of a track; energy captures the intensity or activity level in a song, while duration determines how long a piece lasts. Understanding how these attributes evolve over time can provide insights into the creative trajectories of these artists. This study aims to explore the trends in valence, duration, and energy in the discographies of JJ Lin, Joker Xue, and Leehom Wang, comparing how these elements have changed with the release of new albums over the past two decades.

The remainder of this paper is constructed as follows. @sec-data provides a sample of the downloaded dataset, @sec-res includes three plots comparing Valence, Duration, and Energy in the Music of JJ Lin, Joker Xue, and Leehom Wang, with associated observations and @sec-disc mentions a brief discussion of the result obtained from the graph.

The data is downloaded using R @citeR and Spotify API @spotifyr with the help of usethis package @usethis, and the analysis is conducted with here @here, tidyverse @tidyverse, ggplot2 @ggplot and knitr @knitr.

# Data {#sec-data}
@tbl-sample is a sample of the downloaded data from @spotify_api_sheeran, with relevant columns selected for later analysis(`valence`,`duration_ms`,`energy`).
```{r}
#| echo: false
#| eval: true
#| label: tbl-sample
#| tbl-cap: "Sample Data"
#| warning: false

# Bind the data for all three artists and add an artist_name column
combined_data <- bind_rows(
  jj_lin %>% mutate(artist_name = "JJ Lin"),
  leehom_wang %>% mutate(artist_name = "Leehom Wang"),
  joker_xue %>% mutate(artist_name = "Joker Xue")
)

# Select relevant columns and get the first 5 rows
summary_stats <- combined_data %>%
  select("album_release_date", "valence", "duration_ms", "energy") %>%
  slice(1:5)  # or you can use head(5) instead of slice(1:5)

# Render the summary table using knitr::kable()
kable(summary_stats)
```
A more detailed representation of the dataset can be found in the appendix. It includes the distribution of valence, duration of songs and energy of three singers respectively.

# Results{#sec-res}
## Valence, Duration and Energy of Three Singers' Songs Over Time
```{r}
#| echo: false
#| eval: true
#| label: fig-valence
#| fig-cap: "Valence of Three Singers' Songs Over Time"
#| warning: false
#| fig-pos: "H"

# Select common columns and bind rows
combined_data <- bind_rows(
  select(jj_lin, artist_name, album_release_date, valence),
  select(leehom_wang, artist_name, album_release_date, valence),
  select(joker_xue, artist_name, album_release_date, valence)
)

# Create plot
combined_data %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(x = album_release_date, y = valence, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Valence",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```


```{r}
#| echo: false
#| eval: true
#| label: fig-duration
#| fig-cap: "Duration of Three Singers' Songs Over Time"
#| warning: false
#| fig-pos: "H"

combined_data_duration <- bind_rows(
  select(jj_lin, artist_name, album_release_date, duration_ms),
  select(leehom_wang, artist_name, album_release_date, duration_ms),
  select(joker_xue, artist_name, album_release_date, duration_ms)
)

combined_data_duration %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(x = album_release_date, y = duration_ms, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Duration (ms)",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```


```{r}
#| echo: false
#| eval: true
#| label: fig-energy
#| fig-cap: "Energy of Three Singers' Songs Over Time"
#| warning: false
#| fig-pos: "H"

# Select common columns and bind rows
combined_data_energy <- bind_rows(
  select(jj_lin, artist_name, album_release_date, energy),
  select(leehom_wang, artist_name, album_release_date, energy),
  select(joker_xue, artist_name, album_release_date, energy)
)

combined_data_energy %>%
  mutate(album_release_date = ymd(album_release_date)) %>%
  ggplot(aes(x = album_release_date, y = energy, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Energy",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")


```
## Observations
The graphs reveal notable trends in the music of JJ Lin, Joker Xue, and Leehom Wang over time. In terms of valence from @fig-valence, all three artists show a slight decline, indicating a shift toward less positive emotional tones in their songs. Joker Xue exhibits the most significant drop in valence, suggesting a move toward more melancholic or neutral themes. In @fig-duration for song duration, JJ Lin’s songs have become slightly longer over time, while Joker Xue and Leehom Wang have maintained consistent lengths. In terms of energy in @fig-energy, both JJ Lin and Joker Xue show a gradual decrease, indicating a shift to less intense, more mellow music. On the other hand, Leehom Wang's energy has increased, suggesting his recent work has become more dynamic and upbeat. Overall, these trends highlight the evolving styles and emotional expressions of the three artists.

# Discussion {#sec-disc}
The comparative analysis of valence, duration, and energy across the discographies of JJ Lin, Joker Xue, and Leehom Wang reveals distinct trends that reflect both personal artistic development and broader shifts in pop music.

- Valence Trends: Across the three artists, valence—representing the emotional tone—generally shows a declining trend over time. JJ Lin’s music maintains a relatively stable, moderate valence but with a slight decline, suggesting a shift towards more emotionally neutral or melancholic compositions. Joker Xue, on the other hand, exhibits a more pronounced decline, reflecting a trend toward lower valence, potentially highlighting the artist's focus on more emotionally complex or darker themes. Leehom Wang also shows a subtle decrease, but with greater variability, indicating a broader emotional range in his discography.

- Duration Trends: Duration, measured in milliseconds, shows relatively stable trends for Joker Xue and Leehom Wang, with no significant variation in song lengths over time. However, JJ Lin shows a slight upward trend in song duration, indicating that his compositions may be growing longer over the years. This could reflect changes in the artist’s production style, possibly exploring more elaborate arrangements or storytelling within songs.

- Energy Trends: Energy, which represents the intensity and activity level in a song, shows diverging trends among the three artists. JJ Lin's energy shows a slight decline, indicating a move toward more mellow or less intense productions. Similarly, Joker Xue also shows a decline, suggesting a preference for slower, perhaps more introspective music in recent years. In contrast, Leehom Wang's energy shows an upward trend, highlighting an increasing intensity in his later works, which may suggest experimentation with more upbeat or dynamic musical elements.

- Broader Implications: These trends provide interesting insights into how individual artists evolve over time and how their music reflects personal or industry-wide changes. The decreasing valence in Joker Xue’s music could indicate a move towards more mature or introspective themes, while Leehom Wang's increasing energy may reflect his embrace of more dynamic and experimental production techniques. JJ Lin, meanwhile, seems to be extending the length of his compositions while maintaining a balanced emotional tone, perhaps catering to a broader audience.

In summary, this study shows that while there are common trends in the emotional content (valence) and song structure (duration), the artists display significant individuality when it comes to the energy in their music. These findings may reflect the diversity of Chinese pop music and how artists adapt their style over time.

\newpage

\appendix

# Appendix {-}

```{r}
#| echo: false
#| eval: true
#| label: fig-appendix
#| fig-cap: "Plot for Valence, Duration and Energy Distribution"
#| warning: false
#| fig-pos: "H"

# Assuming combined_data is already created
combined_data <- bind_rows(
  jj_lin %>% mutate(artist_name = "JJ Lin"),
  leehom_wang %>% mutate(artist_name = "Leehom Wang"),
  joker_xue %>% mutate(artist_name = "Joker Xue")
)

# Plot for Valence Distribution
ggplot(combined_data, aes(x = valence, fill = artist_name)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  labs(
    x = "Valence",
    y = "Density",
    fill = "Artist"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

# Plot for Duration Distribution
ggplot(combined_data, aes(x = duration_ms, fill = artist_name)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  labs(
    x = "Duration (ms)",
    y = "Density",
    fill = "Artist"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

# Plot for Energy Distribution
ggplot(combined_data, aes(x = energy, fill = artist_name)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  labs(
    x = "Energy",
    y = "Density",
    fill = "Artist"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```


\newpage


# References


